---
title: "CF Input Meta Analysis"
output: html_notebook
---

This code will analyze CF input patterns from a Meta-data sheet. To generate Meta-data sheet, create a csv/excel sheet with the following headers:

```
CellDate
CellID
DendriteCategory
CFSteps
Age
Lobule
SulcalCategory
SplitDistance
SplitAngle
FillQuality
Region
```
When you have your database, load your libraries:

```{r Libraries, echo=TRUE}
library(ggplot2)
library(wesanderson)
library(cowplot)
library(magick)
library(plyr)
```

Copy your csv or excel database 

``` {r Shape data}
#shape data for visualization and stats
x<-read.delim(pipe("pbpaste"),header=T)
x$CFStepCat <- as.factor(x$CFSteps)
#x$FillQuality <- as.factor(x$FillQuality)
x$DendriteCategory <- factor(x$DendriteCategory, levels=c("Normative", "Split", "Double"))
x$SulcalCategory <- factor(x$SulcalCategory, levels=c("Sulcus", "Bank", "Gyrus"))
path = "/Users/sebkasl12/Desktop/Plots"
path2 = path #"/Users/sebkasl12/Desktop/NSF GRFP 2019 App"
x_nodouble = x[which(x$DendriteCategory != "Double"),]
x_justsplit = x[which(x$DendriteCategory == "Split"),]
x_justnorm = x[which(x$DendriteCategory == "Normative"),]
x_justdouble = x[which(x$DendriteCategory == "Double"),]
x_no3CF = x_nodouble[which(x_nodouble$CFStepCat != 3),]
```
If you decide to remove poorest quality data
```{r Quality control}
#only high quality cells (above filling quality of 3)
#x <- x[which(x$FillQuality > 2),]
```

``` {r Get demographics}
#to look at CF input distributions by ratio, not absolute number
cells_by_category <- as.data.frame(table(x$DendriteCategory))
total_Norm_cells <- sum(x$DendriteCategory == "Normative")
total_Split_cells <- sum(x$DendriteCategory == "Split")
total_Double_cells <- sum(x$DendriteCategory == "Double")
total_cells <- nrow(x)

multi_Norm_cells <- sum(x$DendriteCategory == "Normative" & x$CFSteps > 1)
multi_Split_cells <- sum(x$DendriteCategory == "Split" & x$CFSteps > 1)
multi_Double_cells <- sum(x$DendriteCategory == "Double" & x$CFSteps > 1)
multi_total <- sum(x$CFSteps > 1)
multi_of_total <- multi_total / total_cells
split_of_total <- total_Split_cells / total_cells
double_of_total <- total_Double_cells / total_cells

ratio_Norm_cells <- multi_Norm_cells / total_Norm_cells
ratio_Split_cells <- multi_Split_cells / total_Split_cells
ratio_Double_cells <- multi_Double_cells / total_Double_cells
multi_total_Norm <- multi_Norm_cells / multi_total
multi_total_Split <- multi_Split_cells / multi_total
multi_total_Double <- multi_Double_cells / multi_total
```

```{r General Fig 1}
###Plot histogram bar plot of CF Step distributions separated by Dendrite Category (Normal, Split, Double) <<APPLICATION FIGURE 1>>
plotA <- ggplot(data=x, aes(x=CFStepCat, fill=DendriteCategory))+
  geom_histogram(stat = "count")+
  facet_grid(~DendriteCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold",family="Times New Roman"),
        strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(legend.position = "none")+
  ylab("# of Cells")+
  xlab("# of CF Inputs")+
  scale_fill_manual(values = c("#C5E7C7", "#60B3BE", "#0A1C6B"))

ggsave(filename = file.path(path,"Fig 1 CF steps cell observation histogram.png"), width = 8, height = 12, units = "cm")

plotA
```

``` {r General Fig 1.5}
#Segregate percentage plots by each dendrite category to visualize relative values
plottotal <- ggplot(data=x, aes(x=CFSteps))+
  geom_bar(aes(y = ((..count..)/sum(..count..))*100))+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold",family="Times New Roman"),
        strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(plot.title = element_text(size=28,face="bold",family="Times New Roman",hjust=0.5))+
  ylab("% of Cells")+
  xlab("CF Inputs")+
  ggtitle("All cells")+
  ylim(0,100)

plotnorm <- ggplot(data=x_justnorm, aes(x=CFSteps))+
  geom_bar(aes(y = ((..count..)/sum(..count..))*100), fill = "#C5E7C7")+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold",family="Times New Roman"),
        strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  theme(plot.title = element_text(size=28,face="bold",family="Times New Roman",hjust=0.5))+
  ylab("% of Cells")+
  xlab("CF Inputs")+
  ggtitle("Normative")+
  ylim(0,100)

plotsplit <- ggplot(data=x_justsplit, aes(x=CFSteps))+
  geom_bar(aes(y = ((..count..)/sum(..count..))*100), fill = "#60B3BE")+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
            axis.title=element_text(size=28,face="bold",family="Times New Roman"),
            strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  theme(plot.title = element_text(size=28,face="bold",family="Times New Roman",hjust=0.5))+
  ylab("% of Cells")+
  xlab("CF Inputs")+
  ggtitle("Split")+
  ylim(0,100)

plotdouble <- ggplot(data=x_justdouble, aes(x=CFSteps))+
  geom_bar(aes(y = ((..count..)/sum(..count..))*100), fill = "#0A1C6B")+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold",family="Times New Roman"),
        strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())+
  theme(plot.title = element_text(size=28,face="bold",family="Times New Roman",hjust=0.5))+
  ylab("% of Cells")+
  xlab("CF Inputs")+
  ggtitle("Double")+
  ylim(0,100)+
  scale_x_discrete(limits=1:3)

ggdraw()+
  draw_plot(plottotal, x=0, y = 0, width = 0.31, height = 1)+
  draw_plot(plotnorm, x=0.33, y = 0, width = 0.21, height = 1)+
  draw_plot(plotsplit, x=0.56, y = 0, width = 0.21, height = 1)+
  draw_plot(plotdouble, x=0.79, y = 0, width = 0.21, height = 1)

ggsave(filename = file.path(path,"Fig 1.5 CF steps normalized percent histogram.png"), width = 25, height = 10, units = "cm")
```

```{r General Fig 2}
###Plot distribution of Split distances per CF step number (if you don't believe in the groups) <<APPLICATION FIGURE 3>>
plot2b <- ggplot(data=x, aes(x=CFStepCat, y=SplitDistance), fill="black", color="black")+
  #geom_point(size=3)+
  geom_boxplot(width=0.2, outlier.shape = NA)+
  geom_jitter(shape=21,width=0.2, size=2.5, stroke=1)+
  #facet_grid(~SulcalCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  #theme(legend.position = c(0.7,0.9),
   #     legend.direction = "horizontal",
    #    legend.text = element_text(family="Times New Roman"))+
  ylab("Split Distance(um)")+
  xlab("# of CF Inputs")+
  #geom_hline(yintercept = 30, size=0.35, linetype='dashed')+
  #geom_hline(yintercept = 25, size=0.35, linetype='dashed')+
  ylim(0,75)

ggsave(filename = file.path(path,"Fig 2 CF steps by split distance no groups color by age.png"), width = 9, height = 11, units = "cm")

plot2b
```

```{r General Fig 3}
###Plot distribution of Split distances per CF step number separated by group (if you believe in the groups) <<APPLICATION FIGURE 2>>
plot2a <- ggplot(data=x_nodouble, aes(x=CFStepCat, y=SplitDistance, fill=DendriteCategory))+
  #geom_point()+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(shape=21,width=0.25, size=2, stroke=1,color="black", fill="white")+
  facet_grid(~DendriteCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=20, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=28,face="bold", family="Times New Roman"),
        strip.text.x = element_text(size = 20, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  ylab("Split Distance(um)")+
  xlab("# of CF Inputs")+
  scale_fill_manual(values = c("#C5E7C7", "#60B3BE", "#0A1C6B"))+
  #geom_hline(yintercept = 30, size=0.35, linetype='dashed')+
  #geom_hline(yintercept = 25, size=0.35, linetype='dashed')+
  ylim(0,75)

ggsave(filename = file.path(path,"Fig 3 CF step and Split distance by dendrite category.png"), width = 6, height = 10, units = "cm")

plot2a
```

```{r General Fig 4}
ggplot(data=x, aes(x=SplitDistance))+
  geom_histogram(binwidth = 10, fill="black", color="white")+
  theme_classic()

ggsave(filename = file.path(path,"Fig 4 PC split distance distribution.png"), width = 5, height = 8, units = "cm")
```

```{r General Fig 4.5}
ggplot(data=x, aes(x=SplitDistance))+
  geom_histogram(binwidth = 15, fill="black", color="white")+
  theme_classic()

ggsave(filename = file.path(path,"Fig 4.5 PC split angle distribution.png"), width = 5, height = 8, units = "cm")
```

```{r General Fig 4.8}
ggplot(data=x_no3CF, aes(x=SplitDistance, group=CFStepCat,fill=CFStepCat))+
  geom_density(alpha=0.6)+
  theme_classic()

ggsave(filename = file.path(path,"Fig 4.8 Split Distance densities by CFs.png"), width = 10, height = 5, units = "cm")

```

```{r General Fig 5}
#Plot distribution of ages per CF step number and dendritic category
ggplot(data=x, aes(x=CFStepCat, y= Age, fill=DendriteCategory))+
  geom_boxplot()+
  geom_jitter(shape=21,width=0.2, size=2, stroke=1,color="black", fill="white")+
  facet_grid(~DendriteCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=18, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=20,face="bold",family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(legend.position = "none")+
  ylab("Age")+
  xlab("# of CF Inputs")+
  ylim(0,70)+
  scale_fill_manual(values = c("#C5E7C7", "#60B3BE", "#0A1C6B"))

ggsave(filename = file.path(path,"Fig 5 Distribution of ages by category and CFstep number.png"), width = 8, height = 8, units = "cm")
```

```{r General Fig 6, warning=FALSE}
#Plot histogram bar plot of CF Step distributions separated by Lobule Region (Gyrus, Bank, Sulcus)
ggplot(data=x, aes(x=CFStepCat, fill=SulcalCategory))+
  geom_histogram(stat = "count")+
  facet_grid(~SulcalCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=18, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=20,face="bold",family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(legend.position = "none")+
  ylab("# of Cells")+
  xlab("# of CF Inputs")+
  scale_fill_manual(values = wes_palette(n=3, name="GrandBudapest1"))

ggsave(filename = file.path(path,"Fig 6 CF steps by lobule region histogram.png"), width = 5, height = 7, units = "cm")
```

```{r General Fig 7, warning=FALSE}
#Plot histogram bar plot of morphological category separated by Lobule Region (Gyrus, Bank, Sulcus)
ggplot(data=x, aes(x=DendriteCategory, fill=SulcalCategory))+
  geom_histogram(stat = "count")+
  facet_grid(~SulcalCategory)+
  theme_classic()+
  theme(axis.text=element_text(size=10, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=20,face="bold",family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(legend.position = "none")+
  ylab("# of Cells")+
  xlab("Cell Morphologies")+
  scale_fill_manual(values = wes_palette(n=3, name="GrandBudapest1"))

ggsave(filename = file.path(path,"Fig 7 Morphologies by lobule region histogram.png"), width = 14, height = 8, units = "cm")

```

To plot binned split distance data, create binned database with these headers:

```
DistanceBin
1CF #num cells in that distance bin with 1CF
2+CF #num cells in that distance bin with 2+CF
Ratio #num 1CF divided by sum of 1CF and 2+CF
Totalcells #sum of 1CF and 2+CF
```

Copy the binned data you want to plot

```{r General Fig 8.5}
#Import binned data from sheet 2 of database
x1<-read.delim(pipe("pbpaste"),header=T)

ggplot(data=x1, aes(x=DistanceBin, y=Ratio, size=Totalcells))+
  geom_point()+
  theme_classic()+
  theme(axis.text=element_text(size=10, face = "bold", family="Times New Roman"),
        axis.title=element_text(size=20,face="bold",family="Times New Roman"),
        strip.text.x = element_text(size = 10, face = "bold", family="Times New Roman"))+
  theme(legend.position = "none")+
  theme(plot.title = element_text(size=20,face="bold",family="Times New Roman",hjust=0.5))+
  ylab("Ratio of cells with 1CF")+
  xlab("Split Distance Bins (um)")+
  geom_hline(yintercept = 1, size=0.35, linetype='dashed')+
  ylim(0,1.3)

ggsave(filename = file.path(path,"Fig 8.5 Ratio 1CF cells by Split Distance.png"), width = 10,height=10, units = "cm")

```

```{r General Fig 9 plots, warning=FALSE, include=FALSE}
#To plot 2D histogram with point graph of Split distance by Split Angle
multiplot1 <- ggplot(data=x_no3CF, aes(x=SplitDistance, y=SplitAngle,color=CFStepCat))+
  geom_point(size=3)+
  theme_classic()+
  scale_y_continuous(position = "right")+
  theme(axis.title = element_blank())+
  theme(legend.position = "none")+
  scale_color_manual(values = c("#e8d73a", "#a33658"))+
  geom_vline(xintercept = 40, size=0.35, linetype="dashed")

multiplot2 <- ggplot(data=x_no3CF, aes(x=SplitDistance, group=CFStepCat,fill=CFStepCat))+
  geom_density(alpha=0.6)+
  theme_classic()+
  theme(axis.title = element_blank())+
  theme(axis.text.y = element_blank())+
  labs(fill = "#CFs")+
  theme(legend.position = "bottom")+
  scale_y_reverse(position = "right")+
  scale_fill_manual(values = c("#e8d73a", "#a33658"))

multiplot3 <- ggplot(data=x_no3CF, aes(x=SplitDistance, group=DendriteCategory,fill=DendriteCategory))+
  geom_density(alpha=0.6)+
  theme_classic()+
  theme(axis.title = element_blank())+
  theme(axis.text.y = element_blank())+
  theme(legend.title = element_blank())+
  scale_y_continuous(position = "right")+
  theme(legend.position = "bottom")+
  scale_y_reverse(position = "right")+
  scale_fill_manual(values = c("#177804", "#032896"))

multiplot4 <- ggplot(data=x_no3CF, aes(x=SplitAngle, group=CFStepCat,fill=CFStepCat))+
  geom_density(alpha=0.6)+
  theme_classic()+
  theme(axis.title = element_blank())+
  theme(axis.text.x = element_blank())+
  labs(fill = "#CFs")+
  theme(legend.position = "right")+
  coord_flip()+
  scale_fill_manual(values = c("#e8d73a", "#a33658"))

multiplot5 <- ggplot(data=x_no3CF, aes(x=SplitAngle, group=DendriteCategory,fill=DendriteCategory))+
  geom_density(alpha=0.6)+
  theme_classic()+
  theme(axis.title = element_blank())+
  theme(axis.text.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.position = "right")+
  coord_flip()+
  scale_fill_manual(values = c("#177804", "#032896"))
```

```{r General Fig 9, warning=FALSE}
ggdraw()+
  draw_plot(multiplot1, x=0, y = 0.5, width = 0.5, height = 0.5)+
  draw_plot(multiplot2, x=0, y = 0.25, width = 0.475, height = 0.25)+
  draw_plot(multiplot3, x=0, y = 0, width = 0.475, height = 0.25)+
  draw_plot(multiplot4, x=0.5, y = 0.523, width = 0.25, height = 0.477)+
  draw_plot(multiplot5, x=0.75, y = 0.523, width = 0.25, height = 0.477)
  
ggsave(filename = file.path(path,"Fig 9 2D hist Step+morph densities by split dist+angle.png"), width = 20, height = 20, units = "cm")
```

```{r Load images for fig 0}
setwd("~/Desktop/Plots")
image <- image_read('General Plots Norm Cell Example.png')
plot1a <- image_ggplot(image_scale(image, "100"), interpolate = FALSE)
image <- image_read('General Plots Split Cell Example.png')
plot1b <- image_ggplot(image_scale(image, "100"), interpolate = FALSE)
image <- image_read('General Plots Double Cell Example.png')
plot1c <- image_ggplot(image_scale(image, "100"), interpolate = FALSE)
```

```{r General Fig 0, warning=FALSE}
setwd("~/Desktop")
plot1 <- ggdraw()+
  draw_plot(plot1a, x=0, y = 0.15, width = 0.33, height = 0.8)+
  draw_plot(plot1b, x=0.345, y = 0.15, width = 0.33, height = 0.8)+
  draw_plot(plot1c, x=0.69, y = 0.15, width = 0.33, height = 0.8)
  
#plotleft <-plot_grid(plot1, plotA, ncol = 1, nrow = 2, labels = c("A"), rel.heights = c(1,3))
plot2 <- plot_grid(plot2b,plot2a, ncol = 1,nrow = 2)
#plot_grid(plotA, plot2, ncol=2, nrow = 1, labels = c("A",""), label_size = 20)

ggdraw()+
  draw_plot(plotA, x=0, y = 0, width = 0.65, height = 0.73)+
  draw_plot(plot2, x=0.65, y = 0, width = 0.35, height = 1)+
  draw_plot(plot1, x=0.07, y = 0.64, width = .56, height = 0.4)

ggsave(filename = file.path(path2,"GRFP App Fig1 .png"), width = 25, height = 20, units = "cm")
```

```{r}
#Plot distribution of cells by lobule per CF step number and dendritic category 
# x$LobuleName <- as.factor(x$Lobule)
# 
# ggplot(data=x, aes(x=CFStepCat, fill = DendriteCategory))+
#   geom_histogram(stat="count")+
#   facet_grid(~LobuleName)+
#   theme_classic()+
#   theme(legend.position = "none")+
#   ylab("Number of Cells")+
#   xlab("Number of CF Inputs")
```

```{r Run Statistics}
#Stats
#Create count table of observations per category of independent and dependent variable
tbl <- table(x$DendriteCategory,x$CFSteps)
#Use Fisher Exact test (small sample size version of a Chi sqared test?)
fisher.test(tbl)

#Above by dendrite category, below stats for other factors (Age, Lobule)
tbl <- table(x_nodouble$SplitDistance,x_nodouble$CFSteps)
fisher.test(tbl)
```



















